
---
title: "HW_11"
author: "Tobi Ore & Luis Andres Guillen"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(kableExtra)
library(tidyverse)
library(knitr)
```

# Question 1:

+ Were rates of survival higher among children or adults?
```{r}
survival_age <- data.frame(apply(Titanic, c(3, 4), sum))
survival_rate <- survival_age$Yes/(survival_age$Yes + survival_age$No)
survival_age <- cbind(survival_age,survival_rate)
answer <- rownames(survival_age)[which.max(survival_age$survival_rate)]
sprintf("The rate of survival is higher among the %s age group",answer)
```

+ Were rates of survival higher among men or women?
```{r}
survival_sex <- data.frame(apply(Titanic, c(2, 4), sum))
survival_rate <- survival_sex$Yes/(survival_sex$Yes + survival_sex$No)
survival_sex <- cbind(survival_sex,survival_rate)
answer <- rownames(survival_sex)[which.max(survival_sex$survival_rate)]
sprintf("The rate of survival is higher among the %s sex group",answer)
```
Alternative solution:

In order to solve Q1. we used the `apply` function called `aggregate`:
```{r}
data(Titanic)
Titanic<-as.data.frame(Titanic)
Survival_age<-aggregate(Freq~Age+Survived,data=Titanic,sum)
children_rate<-Survival_age$Freq[3]/(Survival_age$Freq[3]+Survival_age$Freq[1])
adult_rate<-Survival_age$Freq[4]/(Survival_age$Freq[2]+Survival_age$Freq[4])
Survival_sex<-aggregate(Freq~Sex+Survived,data=Titanic,sum)
men_rate<-Survival_sex$Freq[3]/(Survival_sex$Freq[3]+Survival_sex$Freq[1])
women_rate<-Survival_sex$Freq[4]/(Survival_sex$Freq[2]+Survival_sex$Freq[4])

```
Based on the analysis, we can conclude that, as expected, children (`r round(children_rate,2)`) and women (`r round(women_rate,2)`) had higher survival rates than adults (`r round(adult_rate,2)`) and men (`r round(men_rate,2)`). 


# Question 2:

Q2 was done in two steps. First, we created the list of data frames:
``` {r } 
prismDat<-function(path){
  data.path<-path
  prism.files<-list.files(path)
  file.names<-sub("cv_","", prism.files)
  file.names<-sub("_daily.csv","", file.names)
  glob.path <- paste0(data.path, "*", ".csv")  # this is the englobing string that can find the files. 
  dataFiles <- lapply(Sys.glob(glob.path), read.csv, skip=11, head=F)
  climvars <- c("pdate", "ppt", "tmin", "tmean", "tmax")
  dataFiles <- lapply(dataFiles, setNames, climvars)
  names(dataFiles)<-file.names
  return(dataFiles)
}

mydata<-prismDat("data/") # retrieving the data

```

Then, we made a function to find the TRUE/FALSE for a given threshold temperature:
``` {r }
temp_check <- function(data, thresh){
  data <- read.csv("./data/cv_grid1_daily.csv", col.names = c("pdate", "ppt", "tmin", "tmean", "tmax") ,skip = 11, header=F)
  threshold <- data$tmin > thresh
  data <- cbind(data,threshold)
  return(data)
}
```
To utilize the function, we can simply use `lapply`, as follows:
```{r }
test<-lapply(mydata,temp_check,-5) #testing function
knitr::kable(head(test$grid1,2)) %>% kable_styling() #showing results
```

# Question 3:

Q3 requires to load several libraries to the environment. The personal NOAA key also needs to be declared to access the data from NCDC:
```{r, include=FALSE}
mykey<-"iczdfxvFGPnTLpqiOCvbjsIpOhTFMUTi"
```
```{r, message=FALSE}
library(rnoaa)
library(lubridate)
library(dplyr)
options(nooakey = mykey)
```
The next step is to create a loop that retrieves every year's timeseries from NOAA's website. 
```{r, warning=FALSE}
data <- data.frame()
for (year in seq(1985,2019)){
  temp_data <- ncdc(datasetid='GSOM', stationid = 'GHCND:ASN00091245', datatypeid='TAVG', startdate = paste0(year,"-01-01"), enddate = paste0(year,"-12-01"), limit = 100, add_units=TRUE)$data
  data<-rbind(data,temp_data)
}
```

# Question 4:
After we have done that procedure, we can rearrange the data to find the mean temperature and the number of months. We use the function `aggregate`, with a specific margin, in our case the years. 

First alternative:
```{r}
kiwi<-data
kiwi$date<-as.Date(substr(kiwi$date,1,10))
kiwi_result<-round(aggregate(kiwi$value,by=list(year(kiwi$date)),mean),3)
kiwi_count<-aggregate(kiwi$date,by=list(year(kiwi$date)),length)
kiwi_result<-cbind(kiwi_result,kiwi_count$x)
names(kiwi_result)<-c("Year","Tavg","Mons")
knitr::kable(head(kiwi_result),align = 'c',caption = "Mean Annual Temperature for Cape Grimp, NZ ")%>% kable_styling()
```
### Except Cape Grim is in Australia - so Oz!
Second alternative (that also shows the years with missing values):
```{r}
annual_avg_temp <- data.frame()
for (year in seq(1985,2019)){
  tavg <- mean(data[grep(paste0("^",year), data$date),]$value)
  mons <- length(data[grep(paste0("^",year), data$date),]$value)
  new_data <- data.frame("Year" = year, "Tavg" = tavg, "Mons" = mons)
  annual_avg_temp <- rbind(annual_avg_temp, new_data)
}
kable(head(annual_avg_temp)) %>% kable_styling()